Maintenance planning on french military aircraft operations

### 2017-11-15 latex/ROADEF/article_roadef_2018.tex
added modifications to article and to R scripts that makes graph

\author{Franco Peschiera, Alain Haït, Olga Battaïa }

\institute{
ISAE... \\
\email{\{franco.peschiera\}@isae-supaero.fr}
}

@@

@@

    Each $v \in \mathcal{V}$ task needs to be accomplished in a planning horizon divided into $t \in \mathcal{T}$ periods of the same length. Each task has a given starting ($startTIME_v \in T$) an ending ($endTIME_v \in T$) time. The starting and ending times are given and cannot be changed.

    Given that the starting and ending times belong to $\mathcal{T}$, the duration of a task is measured in a certain number of periods. Each task requires (seizes) a certain quantity of resources ($REQ_{vt}$). For resources, see \ref{def:resource}.

    Not all $a \in \mathcal{A}$ resources can be used to accomplish any task: each task has specific requirements that a resource needs to satisfy in order to be a candidate. In other words, for each task $t$, there is a set $CAND_{v} \subset \mathcal{A}$ that lists the resources that can be assigned to that task. A resource can be a candidate for more than one task.

@@

    Resources are what makes possible the completion of tasks. These specific resources suffer from wear and tear and so require a recurrent preventive maintenance (see \ref{def:maint}) in order to be in a correct state. These maintenances need to take place before a given amount of time (we will call it $MAX\_ELPASED$) has passed since the previous maintenance or before the resource has been used for a given amount of time (we will call it $MAX\_USAGE$) since the previous maintenance, whichever happens first. While a resource $a$ has not received the proper maintenance, it cannot be used to satisfy tasks and is considered useless.

    We will call "remaining elapsed time" (or $ret$) to the amount of time that needs to pass in order for the resource to need a maintenance.

    We will call "remaining usage time" (or $rut$) to the amount of time that the resource needs to be used in order for it to need a maintenance.

@@

    The following are the possible $s \in \mathcal{S}$ states for a resource in any given time:

@@

    These maintenances take an amount of time equal to $M\_DURATION_m$ periods and cannot be interrupted. The state of the resource after exiting the maintenance period is of "as good as new" in terms of the two indicators that decide the maintenance (remaining elapsed time and remaining usage time).

    In other words, after receiving maintenance of type $m$, a resource restores its remaining elapsed time and remaining usage time back to their max values $MAX\_ELAPSED$ and $MAX\_USAGE$ respectively.

    The number of resources under maintenance in any given period is limited to a maximum capacity ($CAP_t$) depending on each period.

@@

    We have already explained that even if a resource is not being used, it still needs to have a maintenance after a given amount of time (the elapsed time $elapsedTIME$). In order to avoid this problem, the resource can be put into a "storage" state.

@@

    % \begin{figure}[!ht]
    %     \begin{center}
    %         \includegraphics[height=2cm,clip=true]{roadef_logo.eps}
    %         \caption[Fig]{Logo de la ROADEF}
    %         \label{logoRoadef}
    %     \end{center}
    % \end{figure}

    % \begin{equation}
    % E=MC^2\label{emc}
    % \end{equation}

@@

    \begin{tabular}{p{30mm}lp{90mm}}
    $CAP_t$                 & : & number of resources that can be in maintenance in period $t$. \\
    $CAND_v$                & : & resources that satisfy task's $v$ needs. \\
    $REQ\_H_{vt}$           & : & time consumed of any resource when assigned to task $v$ in period $t$. \\
    $REQ\_R_{vt}$           & : & number or resources to assign to task $v$ in period $t$. \\            
    $MAX\_USAGE$            & : & maximum amount of used time that a resource can have between two maintenances. \\    
    $MAX\_ELAPSED$          & : & maximum amount of elapsed time that a resource can have between two maintenances. \\    
    $M\_DURATION$           & : & maintenance duration, in periods. \\        
    $RUT\_INI_a$            & : & remaining used time for resource $a$ at the beginning of horizon. \\        
    $RET\_INI_a$            & : & remaining elapsed time for resource $a$ at the beginning of horizon. \\
    $STATE\_INI_{at}$       & : & some $s \in \mathcal{S}$ in case the resource $a$ has a fixed state. If not; this is not filled. \\
    \end{tabular}
    \bigskip

@@

    \section{Decision variables}

    \begin{tabular}{p{30mm}lp{90mm}}
        $ret_{at}$      & : & remaining elapsed time for resource $a$ after period $t$ (including t=0) \\  
        $rut_{at}$      & : & remaining usage time for resource $a$ after period $t$ (including t=0) \\  
        $task_{avt}$    & : & 1 if resource $a \in CAND_v$ is assigned task $v$ in period $t$ \\  
        $state_{ast}$   & : & 1 if resource $a$ is assigned state $s$ in period $t$ \\  
        $used_{at}$     & : & time consumed by resource $a$ in period $t$ \\
        $start_{at}$    & : & 1 if resource $a$ starts maintenance in period $t$ \\
        $available$     & : & minimum number of available resources in any period \\
        $maintenance$   & : & maximum number of resources in maintenance in any period \\
    \end{tabular}
    \bigskip

    \begin{align}
        %%%%%%%%%%%%%%%%%%%
        % MAIN VARS
        %%%%%%%%%%%%%%%%%%%
        % maximum capacity:
        & \sum_{a \in \mathcal{A}} state_{ast} \leq CAP_{t}
                & t \in \mathcal{T}, s = 'M' \label{eq:capacity}\\
        % minimum resources
        & \sum_{a \in \mathcal{A}} task_{avt} \geq REQ\_R_{vt}
                & t \in \mathcal{T}, v \in \mathcal{V} \label{eq:taskres}\\
        % max one task per period
        & \sum_{v \in \mathcal{V}} task_{avt} \leq 1
                & t \in \mathcal{T}, a \in \mathcal{A} \label{eq:maxtasks}\\
    \end{align}

@@

    \begin{align}
        %%%%%%%%%%%%%%%%%%%
        % BALANCES
        %%%%%%%%%%%%%%%%%%%
        % definition of consumed resource hours:
        & used_{at} \geq \sum_{v \in \mathcal{V}} task_{avt} \times REQ\_H_{vt}
                & a \in \mathcal{A}, t \in \mathcal{T} \label{eq:usage}\\
        % remaining used time
        & rut_{at} \leq (rut_{at-1} - used_{at}) + (MAX\_USAGE \times start_{at})
                & t \in \mathcal{T}, a \in \mathcal{A} \label{eq:balance1}\\
        % remaining elapsed time. TODO: improve this constraint like the maintenance state duration
        & ret_{at} \leq (ret_{at-1} - 1) + (MAX\_ELAPSED \times start_{at})
                & t \in \mathcal{T}, a \in \mathcal{A} \label{eq:balance2}\\
    \end{align}

@@

    \begin{align}
        %%%%%%%%%%%%%%%%%%%
        % STATES
        %%%%%%%%%%%%%%%%%%%
        % if start maintenance, at least X periods
        & state_{ast_2} \geq start_{at_1} \notag\\
                & \hspace{10mm} (t_1, t_2) \in \mathcal{T}, t_1 \leq t_2 \leq t_1 + M\_DURATION - 1, a \in \mathcal{A}, s = \mq{M} \label{eq:start1}\\
        % if maintenance doesn't start, we cannot have maintenance:
        & \sum_{t_1 \in \mathcal{T}} start_{at_1} \geq state_{ast_2} \notag\\
                & \hspace{10mm} (t_1) \in \mathcal{T}, t_1 \leq t_2 \leq t_1 + M\_DURATION - 1, a \in \mathcal{A}, s = \mq{M} \label{eq:start2}\\
        % if assigned to task: in working state
        & state_{ast} \geq task_{avt} \notag\\
                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = \mq{V} \label{eq:state1}\\
        % only one state. If we take out the 'do nothing' state, this could be \leq
        & \sum_{s \in \mathcal{S}} state_{ast} = 1 \notag\\
                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:state2}\\
    \end{align}

@@

    \begin{align}
        %%%%%%%%%%%%%%%%%%%
        % OBJECTIVE
        %%%%%%%%%%%%%%%%%%%
        % min available
        & available \leq \sum_{a} state_{ast}
                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = \mq{N} \label{eq:available}\\
        % max maintenance
        & maintenance \geq \sum_{a} state_{ast}
                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = \mq{N} \label{eq:maintenance}\\
    \end{align}

@@

    \begin{align}
        %%%%%%%%%%%%%%%%%%%
        % HORIZON
        %%%%%%%%%%%%%%%%%%%
        % TODO: initialize maintenance states too!
        % rut initialize
        & rut_{at} = RUT\_INI_a
                & \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:remaining1}\\
        % ret initialize
        & ret_{at} = RET\_INI_a
                & \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:remaining2}\\
        % fixed states during planning period
        & state_{ast} = (STATE\_INI_{at} = s)
                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}\, | if\, STATE\_INI_{at} \neq \mq{} \label{eq:fixedstates}\\
    \end{align}

@@

\bibliographystyle{plain}

% \begin{thebibliography}{2}
% \bibitem{toth02}
% Paolo Toth and Daniele Vigo.
% \newblock \emph{The Vehicle Routing Problem}.
% \newblock Monographs on Discrete Mathematics and Applications. Society for Industrial and Applied Mathematics, 2002.
% \bibitem{kirkpatrick83}
% Scott Kirkpatrick, C~Daniel Gelatt, and Mario~P Vecchi.
% \newblock Optimization by simmulated annealing.
% \newblock \emph{science}, 220\penalty0 (4598):\penalty0 671--680, 1983.

>>>>

\author{Franco Peschiera\inst{1}, Alain Haït\inst{1}, Olga Battaïa\inst{1}, Nicolas Dupin\inst{2}}

% Les affiliations (par ordre croissant des numéros d'affiliation) séparées par \and
\institute{
Institut Supérieur de l'Aéronautique et de l'Espace \\
\email{\{franco.peschiera,alain.haït,olga.battaia\}@isae-supaero.fr}
\and
Direction Générale de l'Armement \\
\email{nicolas.dupin@intradef.gouv.fr}
}

@@

    Le problème des planifications de maintenances des avions commerciales est un sujet qui a été étudié beaucoup dans la littérature. Dans ce problème on affecte des avions à vols au même temps qu'on les affecte à des maintenances. Il est aussi similaire à la planification des maintenances de matériel roulant.

    Par contre, le problème des planifications de maintenances des avions militaires n'a pas été étudié que pour quelques auteurs. Les principales différences se trouve aux objectives, à la taille de l'horizon de planification et au fait que les avions retourne toujours à la même base après les opérations.

    L'objectif dans le Military Flight and Maintenance Planning n'est pas ce de minimiser le coût total de la planification si non ce de lisser les plus possible les opérations de maintenances au long de l'horizon au même temps que de lisser le plus possible la disponibilité des avions \cite{Kozanidis2008}, \cite{Cho2011} et \cite{Verhoeff2015}.

    En plus, chaque armée est différente une de l'autre et, donc, les contraints et buts qui conditionnent les opérations de maintenances sont aussi différentes.

    On a formulé des nouveaux contraints, comme la possibilité de stocker temporairement des avions ou la nécessité de affecter des avions dans un temps minimal au missions. Ces contraints font que le modèle devient plus complexe en ajoutant une dépendance plus fort entre les mois consécutifs.

@@

    Each $v \in \mathcal{V}$ task needs to be done in a planning horizon divided into $t \in \mathcal{T}$ periods of the same length. Each task has a given starting ($startTIME_v \in T$) an ending ($endTIME_v \in T$) time. The starting and ending times are given and cannot be changed.

    Each task $v$ requires (seizes) a certain quantity of resources $a \in \mathcal{A}$ each period $t$ ($REQ_{vt}$). For resources, see \ref{def:resource}.

    Not all $a$ resources can be used to accomplish any task: each task has specific requirements that a resource needs to satisfy in order to be a candidate. In other words, for each task $v$, there is a set $cd_{v} \subset \mathcal{A}$ that lists the resources that can be assigned to that task. A resource can be a candidate for more than one task.

@@

    Resources are what makes possible the completion of tasks. These specific resources suffer from wear and tear and so require a recurrent preventive maintenance (see \ref{def:maint}) in order to be in a correct state. These maintenances need to take place before a given amount of time (that we will call $MAX\_ELPASED$) has passed since the previous maintenance or before the resource has been used for a given amount of time (that we will call $mu$) since the previous maintenance, whichever happens first. While a resource $a$ has not received the proper maintenance, it cannot be used to satisfy tasks and is considered useless.

    We will call "remaining elapsed time" (or $ret$) to the amount of time that needs to pass in order for the resource to need a maintenance at any given period.

    We will call "remaining usage time" (or $rut$) to the amount of time that the resource needs to be used in order for it to need a maintenance at any given period.

@@

    The following are the possible $s \in \mathcal{S}$ states for a resource to be in any given time:

@@

    These maintenances take an amount of time equal to $md_m$ periods and cannot be interrupted. The state of the resource after exiting the maintenance period is of "as good as new" in terms of the two indicators that decide the maintenance (remaining elapsed time and remaining usage time).

    In other words, after receiving a maintenance, a resource restores its remaining elapsed time and remaining usage time back to their max values $me$ and $mu$ respectively.

    The number of resources under maintenance in any given period is limited to a maximum capacity ($c_t$) which depends on each period.

@@

    We have already explained that even if a resource is not being used, it still needs to have a maintenance after a given amount of time (govern by the elapsed time $elapsedTIME$). In order to avoid this problem, the resource can be put into a "storage" state.

@@

@@

    \begin{tabular}{p{10mm}lp{110mm}}
    $c_t$                 & : & number of resources that can be in maintenance in period $t$. \\
    $cd_v$                & : & resources that satisfy task's $v$ needs. \\
    $rh_{vt}$           & : & time consumed of any resource when assigned to task $v$ in period $t$. \\
    $rr_{vt}$           & : & number or resources to assign to task $v$ in period $t$. \\            
    $mu$            & : & maximum used time for each resource between maintenances. \\    
    $me$          & : & maximum elapsed time for each resource between maintenances. \\    
    $md$           & : & maintenance duration, in periods. \\        
    $rut\_i_a$            & : & remaining used time for resource $a$ at the beginning of horizon. \\        
    $ret\_i_a$            & : & remaining elapsed time for resource $a$ at the beginning of horizon. \\
    $state\_i_{at}$       & : & some $s \in \mathcal{S}$ in case the resource $a$ has a fixed state. \\
    \end{tabular}
    \bigskip

@@

    \begin{tabular}{p{20mm}lp{100mm}}
        $ret_{at}$      & : & remaining elapsed time for resource $a$ after period $t$ \\  
        $rut_{at}$      & : & remaining usage time for resource $a$ after period $t$ \\  
        $task_{avt}$    & : & 1 if resource $a \in cd_v$ is assigned task $v$ in period $t$ \\  
        $state_{ast}$   & : & 1 if resource $a$ is assigned state $s$ in period $t$ \\  
        $used_{at}$     & : & time consumed by resource $a$ in period $t$ \\
        $start_{at}$    & : & 1 if resource $a$ starts maintenance in period $t$ \\
        $available$     & : & minimum number of available resources in any period \\
        $maintenance$   & : & maximum number of resources in maintenance in any period \\
    \end{tabular}
    \bigskip

    \begin{align}
        %%%%%%%%%%%%%%%%%%%
        % MAIN VARS
        %%%%%%%%%%%%%%%%%%%
        % maximum capacity:
        & \sum_{a \in \mathcal{A}} state_{ast} \leq c_{t}
                & t \in \mathcal{T}, s = 'M' \label{eq:capacity}\\
        % minimum resources
        & \sum_{a \in \mathcal{A}} task_{avt} \geq rr_{vt}
                & t \in \mathcal{T}, v \in \mathcal{V} \label{eq:taskres}\\
        % max one task per period
        & \sum_{v \in \mathcal{V}} task_{avt} \leq 1
                & t \in \mathcal{T}, a \in \mathcal{A} \label{eq:maxtasks}
    \end{align}

@@

    \begin{align}
        %%%%%%%%%%%%%%%%%%%
        % BALANCES
        %%%%%%%%%%%%%%%%%%%
        % definition of consumed resource hours:
        & used_{at} \geq \sum_{v \in \mathcal{V}} task_{avt} \times rh_{vt}
                & a \in \mathcal{A}, t \in \mathcal{T} \label{eq:usage}\\
        % remaining used time
        & rut_{at} \leq (rut_{at-1} - used_{at}) + (mu \times start_{at})
                & t \in \mathcal{T}, a \in \mathcal{A} \label{eq:balance1}\\
        % remaining elapsed time. TODO: improve this constraint like the maintenance state duration
        & ret_{at} \leq (ret_{at-1} - 1) + (me \times start_{at})
                & t \in \mathcal{T}, a \in \mathcal{A} \label{eq:balance2}
    \end{align}

@@

    \begin{align}
        %%%%%%%%%%%%%%%%%%%
        % STATES
        %%%%%%%%%%%%%%%%%%%
        % if start maintenance, at least X periods
        & state_{ast_2} \geq start_{at_1} 
                & \hspace{10mm} (t_1, t_2) \in \mathcal{T}, t_1 \leq t_2 \leq t_1 + md - 1, a \in \mathcal{A}, s = \mq{M} \label{eq:start1}\\
        % if maintenance doesn't start, we cannot have maintenance:
        & \sum_{t_1 \in \mathcal{T}} start_{at_1} \geq state_{ast_2} 
                & \hspace{10mm} (t_1) \in \mathcal{T}, t_1 \leq t_2 \leq t_1 + md - 1, a \in \mathcal{A}, s = \mq{M} \label{eq:start2}\\
        % if assigned to task: in working state
        & state_{ast} \geq task_{avt} 
                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = \mq{V} \label{eq:state1}\\
        % only one state. If we take out the 'do nothing' state, this could be \leq
        & \sum_{s \in \mathcal{S}} state_{ast} = 1 
                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:state2}
    \end{align}

@@

    \begin{align}
        %%%%%%%%%%%%%%%%%%%
        % OBJECTIVE
        %%%%%%%%%%%%%%%%%%%
        % min available
        & available \leq \sum_{a} state_{ast}
                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = \mq{N} \label{eq:available}\\
        % max maintenance
        & maintenance \geq \sum_{a} state_{ast}
                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = \mq{N} \label{eq:maintenance}
    \end{align}

@@

    \begin{align}
        %%%%%%%%%%%%%%%%%%%
        % HORIZON
        %%%%%%%%%%%%%%%%%%%
        % TODO: initialize maintenance states too!
        % rut initialize
        & rut_{at} = rut\_ini_a
                & \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:remaining1}\\
        % ret initialize
        & ret_{at} = ret\_ini_a
                & \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:remaining2}\\
        % fixed states during planning period
        & state_{ast} = 1
                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}; s \in \mathcal{S}\, | if\, state\_ini_{at} = s \label{eq:fixedstates}
    \end{align}

@@

\bibliographystyle{plain}
\bibliography{MFMP}
% \begin{thebibliography}{2}
% \bibitem{toth02}
% Paolo Toth and Daniele Vigo.
% \newblock \emph{The Vehicle Routing Problem}.
% \newblock Monographs on Discrete Mathematics and Applications. Society for Industrial and Applied Mathematics, 2002.
% \bibitem{kirkpatrick83}
% Scott Kirkpatrick, C~Daniel Gelatt, and Mario~P Vecchi.
% \newblock Optimization by simmulated annealing.
% \newblock \emph{science}, 220\penalty0 (4598):\penalty0 671--680, 1983.
